# Quarto and R {#sec-r}

When you include R code cells in your document, Quarto uses the knitr engine to execute them.
The basics of code cells, including syntax and common options like `echo` and `include`, are covered in @sec-code.
This chapter covers the features specific to R and knitr.

## Output {#sec-r-output}

### Inline code {#sec-inline-code}

In addition to code cells, you can embed R code directly within your markdown text using inline code, inserting computed values into your narrative without manually copying them.

The syntax uses single backticks with `{r}` immediately after the opening backtick:

```markdown
The mean temperature was `{{r}} mean(airquality$Temp)` degrees.
```

We recommend that you avoid complex function calls or operations that do significant work in inline code and instead compute values in a code cell and then reference them inline:

```{r}
avg_temp <- mean(airquality$Temp)
```

```markdown
The mean temperature was `{{r}} avg_temp` degrees.
```

This renders as: The mean temperature was `r avg_temp` degrees.

By default, inline code produces plain text.
To generate markdown-formatted output, wrap the result in `I()`:

```markdown
The result is `{{r}} I("**significant**")`.
```

This produces bold text in the rendered output.

::: callout-note
## Inline code in YAML

Inline code cannot be used directly in YAML strings, such as titles or figure captions.
For dynamic captions, use the cross-reference div syntax instead (see @sec-crossreferences).
:::

### Data frame printing {#sec-df-print}

Quarto provides the `df-print` option to control how data frames appear in your output, and it can be set at the document-level in the YAML header:

```yaml
---
title: My Analysis
format:
  html:
    df-print: paged
---
```

The `default` option uses the standard S3 print method, i.e., tables appear as they would in the R console.
The `tibble` option prints the table as a plain text tibble, which is equivalent to wrapping the resulting table with `tibble::tibble()`.
The `kable` option applies `knitr::kable()` and supports various output formats, including HTML, PDF, and Word.
In HTML documents, the `paged` option enables interactive tables with pagination.

### Graphics devices {#sec-graphics-devices}

When R generates a plot, knitr uses a *graphics device* to save it to a file, which is then included in the output.
The default device is `png` for HTML formats, `pdf` for LaTeX, and `svg` for typst.
You may want to change it to improve text rendering, achieve scalable output, or ensure PDF compatibility.

You can set the graphics device with the `dev` option, either at the document level or for individual cells:

```yaml
---
knitr:
  opts_chunk:
    dev: ragg_png
---
```

Or for a single cell:

```{{r}}
#| dev: svglite
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```

Common choices include the following:

- The `png` device is the default.
- The `ragg_png` device creates higher-quality PNG images for better text rendering and anti-aliasing. Generating figures with this device requires the [**ragg**](https://ragg.r-lib.org/) package.
- The `svglite` device creates scalable HTML output in vector format. Figures created with this device are sharp at any zoom level. This device requires the [**svglite**](https://svglite.r-lib.org/) package.
- The `pdf` device generates figures in standard PDF format. It is the default vector device for LaTeX/PDF output.
- The `cairo_pdf` device generates figures in PDF format with Unicode text. It is especially useful for handling non-Latin characters and special symbols.

::: callout-tip
## Automatic device selection for PDF

The R option `knitr.graphics.auto_pdf` (shown in @sec-execution-options-r) tells knitr to automatically use a PDF graphics device when rendering to PDF output and a PNG device otherwise.
That way, you don't need to set `dev` differently for each format.
:::

While not knitr-specific, the `fig-format` option is also worth knowing about alongside `dev`.
It works across all Quarto engines:

```yaml
---
format:
  html:
    fig-format: svg
---
```

Use this when you only need to change the output file format without customizing the specific graphics device.
When both `dev` and `fig-format` are set, `dev` takes precedence.

### Interactive output {#sec-interactive-output-r}

Many R packages produce interactive HTML output---plots, tables, maps---through the [htmlwidgets](https://www.htmlwidgets.org/) framework.
Load the package and call the function as you would in an interactive R session.
For example, you can use code cells that produce interactive plots with [plotly](https://plotly.com/r/):

```{{r}}
#| eval: false
p <- ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point()

ggplotly(p)
```

Or interactive tables with [DT](https://rstudio.github.io/DT/):

```{{r}}
#| eval: false
DT::datatable(mtcars, options = list(pageLength = 5))
```

Or interactive maps with [leaflet](https://rstudio.github.io/leaflet/):

```{{r}}
#| eval: false
leaflet() |>
  addTiles() |>
  addMarkers(lng = -73.99, lat = 40.73, popup = "New York City")
```

It is important to note that interactive widgets only work in HTML output formats.
When rendering to PDF or Word, Quarto defers to knitr's behavior, which by default is to stop with an error reporting that functions producing HTML output were found in a document targeting a non-HTML format.

There are two ways around this.
The first is to install the [webshot2](https://rstudio.github.io/webshot2/) package along with Google Chrome or another Chromium-based browser.
When both are available, knitr automatically takes a static screenshot of each widget and includes that image in the output instead:

```r
install.packages("webshot2")
```

The second is to use `knitr::is_html_output()` (see @sec-conditional-execution) to conditionally include widgets only in HTML output, falling back to a static plot or table for other formats.
This gives you more control over the non-HTML version than a screenshot does, at the cost of writing the fallback yourself.

::: callout-note
## Widget dependencies and self-contained documents

Each HTML widget brings its own JavaScript and CSS dependencies.
In self-contained HTML documents (`embed-resources: true`), these dependencies are bundled into the file, which can increase its size.
If your document uses many different widget packages, consider using the default non-self-contained output and hosting the result as a directory.
:::

### Raw markdown output

Sometimes you want your R code to produce raw markdown that Quarto processes directly, rather than being wrapped in output containers.
Use `output: asis` for this, and you will usually want `echo: false` as well:

```{{r}}
#| output: asis
#| echo: false
cat("## A Dynamic Heading\n\n")
cat("This paragraph was generated by R.\n")
```

One common use case is programmatically generating sections in a loop.
For example, you can iterate over the groups in a dataset, emitting a heading and some content for each:

```{{r}}
#| output: asis
#| echo: false
for (cyl in sort(unique(mtcars$cyl))) {
  cat("### Cars with", cyl, "cylinders\n\n")

  group <- mtcars[mtcars$cyl == cyl, ]
  cat("Mean fuel efficiency:", round(mean(group$mpg), 1), "mpg.\n\n")
}
```

Separating each piece of output with `\n\n` is important: the blank line tells Quarto to treat the heading and the paragraph as distinct markdown blocks rather than running them together.
For example, the code cell above results in the following output:

::: {.rendered}
```{r}
#| echo: false
#| output: asis
for (cyl in sort(unique(mtcars$cyl))) {
  cat("### Cars with", cyl, "cylinders {.unnumbered .unlisted}\n\n")

  group <- mtcars[mtcars$cyl == cyl, ]
  cat("Mean fuel efficiency:", round(mean(group$mpg), 1), "mpg.\n\n")
}
```
:::

This approach works well for text output, but not for code cells that produce figures or tables.[^r-1]

[^r-1]: For situations like this we refer readers to the R Markdown Cookbook [@xie2020] for knitr functionality on [child documents](https://pkg.yihui.org/rmarkdown-cookbook/child-document) and [`knit_expand()`](https://pkg.yihui.org/rmarkdown-cookbook/knit-expand).

## Execution {#sec-execution-options-r}

@sec-code introduced cell-level options like `echo` and `include`.
Many of these options can also be set at the document level to apply to all code cells.

### Document-level execution options

Set execution options for the entire document using the `execute` key in your YAML header:

```yaml
---
title: My Report
execute:
  echo: false
  warning: false
---
```

Common document-level execution options include the following, which take `true` or `false` unless noted otherwise:

- `eval`: Whether to execute code cells
- `echo`: Whether to display source code
- `output`: Whether to include results (can also be set to `asis`)
- `warning`: Whether to include warnings
- `error`: Whether to include errors (without halting)
- `include`: Whether to include any output at all

Additionally, `fig-width` and `fig-height` can be set at the document level to define default figure dimensions in inches.

Furthermore, the knitr engine has additional options that you can set at the document level using the `knitr` key.
These map directly to knitr's chunk options:

```yaml
---
title: My Document
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    R.options:
      knitr.graphics.auto_pdf: true
---
```

`R.options` temporarily sets R options via `options()` before each code cell executes.
The original values are restored afterward.

### Cell-level execution options

Individual code cells can override document-level settings.
The knitr engine supports a large set of cell-level options.

- Some options control code appearance:
  - `code-fold`: Collapse code into an expandable section
  - `code-summary`: Text shown for collapsed code (e.g., "Show code")
  - `code-line-numbers`: Add line numbers to code

- Some control the output:
  - `output`: Whether and how to show results (`true`, `false`, or `asis` to indicate that the output is raw markdown and should not have any of Quarto's standard enclosing markdown)
  - `results`: How to display text results (`markup`, `asis`, `hold`, `hide`)
  - `collapse`: Merge code and output into a single block
  - `comment`: Prefix for output lines (default `"##"`)

- Some are specific to code cells that generate figures:
  - `fig-cap`: Caption
  - `fig-alt`: Alternative text
  - `fig-align`: Alignment: `left`, `center`, or `right`
  - `fig-dpi`: Resolution in dots per inch
  - `fig-width` and `fig-height`: Figure dimensions in inches
  - `out-width` and `out-height`: Output dimensions as percentage (80%), pixels (400px), or page metrics (0.5\linewidth)

- Finally, some relate to caching:
  - `cache`: Enable caching for the cell
  - `dependson`: Specify cells this one depends on

For a complete list of knitr cell options, see the [Quarto reference](https://quarto.org/docs/reference/cells/cells-knitr.html).

### Conditional execution {#sec-conditional-execution}

Sometimes you need code to behave differently depending on the output format.
knitr provides helper functions that report the format being rendered:

```{{r}}
if (knitr::is_html_output()) {
  # Code for HTML output only
}

if (knitr::is_latex_output()) {
  # Code for PDF output only
}
```

For formats without a dedicated helper, `knitr::pandoc_to()` returns the Pandoc output format, or tests against one when you pass a name:

``` r
knitr::pandoc_to()
# Returns e.g., "html", "latex", "docx"
knitr::pandoc_to("typst")
# Returns TRUE if the output is typst
```

Quarto also exposes format information through the `QUARTO_EXECUTE_INFO` environment variable, which it sets during rendering:[^r-2]

[^r-2]: For more information, see Quarto documentation on execution context information at <https://quarto.org/docs/advanced/quarto-execute-info.html>.

```{{r}}
info <- jsonlite::fromJSON(Sys.getenv("QUARTO_EXECUTE_INFO"))
format_name <- info$format$identifier$`base-format`

if (format_name == "html") {
  # HTML-specific code
}
```

This provides finer-grained format details than `knitr::pandoc_to()`, such as sub-format identifiers and format options.

## Parameters {#sec-parameters}

Parameters let you create variations of a document without editing the source---for example, generating the same report for different regions, time periods, or scenarios.

You can define parameters in your YAML header using the `params` key:

```yaml
---
title: Regional Sales Report
params:
  region: North
  year: 2026
  show_details: true
---
```

Parameters can be accessed in code through the `params` list:

```{{r}}
#| echo: false
sales_data |>
  filter(region == params$region, year == params$year)
```

```markdown
This report covers the `{{r}} params$region` region for `{{r}} params$year`.
```

It is also possible to override parameters from the command line using the `-P` flag:

```bash
quarto render report.qmd -P region:South -P year:2025
```

For multiple parameter sets, you can create a YAML file with the values:

```{.yaml filename="params-south.yml"}
region: South
year: 2025
show_details: false
```

Then reference it when rendering:

```bash
quarto render report.qmd --execute-params params-south.yml
```

::: callout-important
## Parameter values must be valid YAML

However you set them, parameters reach your document through YAML serialization.
Every value must therefore be something YAML can represent: strings, numbers, booleans, and lists or nested maps of those.
You cannot pass R objects as parameters---a data frame, a fitted model, or a function will not survive the trip.

When a parameter needs to stand in for a complex object, pass a value that identifies it instead: a file path, a database query, or an ID that the document uses to load the object itself.
:::

## Rendering {#sec-r-rendering}

### Rendering from R {#sec-render-from-r}

The [**quarto**](https://quarto-dev.github.io/quarto-r/) R package provides functions to render documents directly from R, without using the terminal.
This package is optional---you don't need it to use Quarto with R.
It's a convenience for those who prefer working entirely within the R console.

::: callout-tip
## More than rendering

The **quarto** package also includes helpers for other tasks you might otherwise do by hand:

- Theme helpers such as `theme_brand_ggplot2()` and `theme_brand_gt()` style plots and tables to match the fonts and colors in your project's `_brand.yml` (see @sec-typst-fonts-colors).
- Table helpers such as `tbl_qmd_div()` and `tbl_qmd_span()` let tables built in R code cells contain Quarto markdown, which Quarto then renders correctly in both HTML and LaTeX output.
- YAML helpers such as `write_yaml_metadata_block()` pass computed values from R into document metadata for Pandoc to use, while `yaml_quote_string()` quotes strings so they survive YAML parsing intact.

See the [package reference](https://quarto-dev.github.io/quarto-r/reference/index.html) for the full list.
:::

You can use `quarto_render()` to render a document:

```r
library(quarto)

quarto_render("report.qmd")
quarto_render("report.qmd", output_format = "pdf")
```

For iterative work, `quarto_preview()` starts a live preview that updates as you save changes:

```r
quarto_preview("report.qmd")

# When done:
quarto_preview_stop()
```

You can also pass parameters directly when rendering with `quarto_render()` using the `execute_params` argument:

```r
quarto_render(
  "report.qmd",
  execute_params = list(region = "South", year = 2025)
)
```

Values passed this way are serialized to YAML before reaching your document, so the restrictions in @sec-parameters apply---`execute_params` takes an R list, but not arbitrary R objects:

```r
# Fails: a data frame is not a YAML value
quarto_render("report.qmd", execute_params = list(data = sales_data))
```

### Rendering R scripts {#sec-render-scripts}

You can render R scripts (`.R` files) directly with Quarto, without converting them to `.qmd` files first.
This uses syntax based on `knitr::spin()`.

- Lines starting with `#'` are treated as markdown
- Lines starting with `#|` are code cell options
- All other lines are R code

Here's an example script that includes both markdown and code cells:

```{.r filename="analysis.R"}
#' ---
#' title: Penguin Analysis
#' author: Soraya Drake
#' format: html
#' ---

#' ## Setup

library(palmerpenguins)
library(ggplot2)

#' ## Bill Size Analysis
#'
#' @fig-bill-sizes shows the relationship between
#' bill length and depth.

#| label: fig-bill-sizes
#| fig-cap: Bill dimensions by species
#| warning: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  theme_minimal()
```

You can render the script just like a `.qmd` file:

```bash
quarto render analysis.R
quarto preview analysis.R
```

This lets you keep your analysis as a plain R script that can also be sourced or run interactively.

## Environment {#sec-r-environment}

### How Quarto finds R {#sec-quarto-finds-r}

Quarto looks for an R installation in the following order:

1. The `QUARTO_R` environment variable, if set
2. The `R_HOME` environment variable, if set
3. The R executable found on your system `PATH`
4. On Windows only, the R installation recorded in the registry, and failing that, an `R` directory under `%programfiles%`

On macOS and Linux there are no further fallbacks: if none of the first three locate R, Quarto tries to run `R` directly and the render fails if your shell can't resolve it.

If Quarto is using a different R than you expect, you can override the lookup by setting `QUARTO_R` to the path of your preferred R binary.
You can give it either the binary itself or the directory containing it, and Quarto warns you if the path doesn't exist rather than silently moving on.
In a Quarto project, one way to do this is an [`_environment` file](https://quarto.org/docs/projects/environment.html) alongside `_quarto.yml`:

```{.bash filename="_environment"}
QUARTO_R=/opt/R/4.4.0/bin/R
```

Note that Quarto invokes `Rscript` rather than `R`, and it resolves that name against the directory your path points to.
The setting above therefore sends Quarto to `/opt/R/4.4.0/bin/Rscript`: only the directory matters, not the binary name you happen to write.

Quarto reads `_environment` from the project root during rendering, so this applies to every document in the project.
Use `_environment.local` for values specific to your own machine---Quarto adds `/_*.local` to `.gitignore` to keep them out of version control.
Variables already set in your shell take precedence over the file, and because this is a project-level feature, a standalone document outside a project needs `QUARTO_R` set in your environment instead.

To see which R installation Quarto is currently using, run `quarto check` from the terminal:

```bash
quarto check knitr
```

The output will include the detected R version and path, along with the knitr and rmarkdown package versions, e.g.:

```bash
Quarto 1.10.3
[✓] Checking R installation...........OK
      Version: 4.6.0
      Path: /Library/Frameworks/R.framework/Resources
      LibPaths:
        - /Users/mine/Library/R/arm64/4.6/library
        - /Library/Frameworks/R.framework/Versions/4.6/Resources/library
      knitr: 1.51
      rmarkdown: 2.31

[✓] Checking Knitr engine render......OK
```

If the R path shown doesn't match what you expect, that's likely the source of any "package not found" or version mismatch issues.

::: callout-tip
## Rendering in Positron

Positron can discover multiple R installations on your system and lets you switch between them.
When you render a Quarto document, Positron sets `QUARTO_R` to the R version active in your session, so rendering uses the same R you're working with interactively.
To render with a different version, switch your active R interpreter in Positron before rendering.
:::

### Diagnosing environment issues {#sec-diagnosing-env}

When Quarto renders a document, it starts a fresh R session.
This session might use a different R installation or library paths than what you use interactively.
In particular, packages you've installed in your interactive session may not be available if the rendering session uses different library paths, and environment variables like `PATH` may also differ.

Add a code cell to your document that outputs your session information:

```{{r}}
sessionInfo()
.libPaths()
```

This will show the R version and the library paths used during rendering.
For more detailed diagnostics, use the [**sessioninfo**](https://sessioninfo.r-lib.org/) package:

``` r
sessioninfo::session_info(to_file = "quarto-session-info.txt")
```

Compare either output with your interactive R session to identify discrepancies.
If library paths differ, that explains missing packages; if the R version differs, see @sec-quarto-finds-r for how to control which R installation Quarto uses.

When the R version is the problem, you can watch Quarto perform the search described in @sec-quarto-finds-r by adding `--log-level debug` to `quarto check`:

```bash
quarto check knitr --log-level debug
```

The output names each location Quarto tries, in order, and which one succeeded:

```bash
Quarto version: 1.10.18
Quarto 1.10.18
-- Searching for R binary --
Looking for 'Rscript' in QUARTO_R: undefined
Looking for 'Rscript' in R_HOME: undefined
Looking for 'Rscript' in PATH.
[execProcess] which Rscript
[execProcess] Success: true, code: 0
Found in PATH at /usr/local/bin/Rscript
```

This is the quickest way to tell whether a stale `QUARTO_R` or `R_HOME` is redirecting Quarto to an R installation you didn't intend.

## Multiple languages {#sec-multiple-languages}

While this chapter focuses on R, documents using the knitr engine can include code cells in other languages.
Quarto itself does nothing special here: any document with at least one `{r}` cell is run through the knitr engine, and knitr then evaluates each non-R cell with one of its own language engines---Python, Julia, SQL, and others.

For Python cells, knitr uses the [**reticulate**](https://rstudio.github.io/reticulate/) package by default, which runs R and Python in a single session.
That means values move between the two languages: Python reads R objects through `r.`, and R reads Python objects through `py$`.

```{{r}}
x <- 42
```

```{{python}}
# reticulate makes R objects available through `r.`
y = r.x + 1
```

```{{r}}
# and Python objects through `py$`, once reticulate is attached
library(reticulate)
py$y
```

Because reticulate is the default rather than an add-on, it needs to be installed---if it isn't, the render stops with an error telling you to install it or opt out.
Using it also means Python output is handled properly, with matplotlib figures and pandas data frames rendering as you'd expect [@xie2020].

::: callout-caution
## Opting out of reticulate

Setting the `python.reticulate` knitr option to `false` runs Python cells with the `python` executable on your `PATH` instead:

```yaml
---
knitr:
  opts_chunk:
    python.reticulate: false
---
```

R and Python are then fully independent: `r.` and `py$` are unavailable, and you'd have to pass data between the languages by writing it to files (e.g., CSV or Parquet).
We don't recommend this, as knitr's support for Python is built around reticulate and much of it doesn't apply without it.
:::

For the full set of rules Quarto uses to choose an engine---including what happens in documents with no R cells at all---see @sec-engine-selection.
