---
echo: false
---

```{r}
{{< include ../_code-snippets/setup.R >}}
```

:::{#basic-cell-options-table}

```{r}
#| label: tbl-species-count
#| tbl-cap: "Number of penguins sampled from each species in the Palmer Archipelago"

{{< include ../_code-snippets/species-count.R >}}
```
:::

:::{#basic-cell-options-figure}

```{r}
#| label: fig-scatter
#| fig-cap: "Penguin bill length versus depth by species"
#| fig-alt: "Scatter plot of bill length vs. depth showing distinct species clustering."

{{< include ../_code-snippets/bill-scatter.R >}}
```

:::


::: {#cross-references-link}

Most penguins were Adelie (@tbl-species-count). @fig-scatter shows quite distinct clustering of the three species in bill length versus depth.

:::

## Div Syntax

::: {#fig-scatter}
```{r}
#| fig-alt: "Scatter plot of bill length vs. depth showing distinct species clustering."

{{< include ../_code-snippets/bill-scatter.R >}}
```

Penguin bill length versus depth by species

:::

## Dynamic Captions


::: {#fig-scatter}
```{r}
#| fig-alt: "Scatter plot of bill length vs. depth showing distinct species clustering."

{{< include ../_code-snippets/bill-scatter.R >}}
```

Penguin bill length versus depth by species for `{r} nrow(penguins)` penguins.

:::

## Long Captions

::: {#fig-scatter}

```{r}
#| fig-alt: "Scatter plot of bill length vs. depth showing distinct species clustering."

{{< include ../_code-snippets/bill-scatter.R >}}
```

Scatter plot depicting the relationship between bill length (mm) and bill depth (mm) across different penguin species. 
The clustering of the species reveals morphological differences in bill dimensions: 
**Adelie** penguins have short, deep bills,
whereas **Gentoo** penguins have long, shallow bills.
**Chinstrap** penguins have similar depth bills to Adelie penguins but there bills are longer.

:::